---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
#imports
library(dplyr)
library(tidyverse)
```

```{r}
deaths <- av %>%
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") %>%
  mutate(Time = parse_number(gsub("Death", "", Time)))

head(deaths)
```

Similarly, deal with the returns of characters.

```{r}
# Replace columns for returns
returns <- av %>%
  pivot_longer(cols = starts_with("Return"),
               names_to = "Time",
               values_to = "Return") %>%
  mutate(Time = parse_number(gsub("Return", "", Time)))

head(returns)
```
Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Drop the "Death" and "Return" columns before combining the datasets
deaths <- deaths %>% select(-starts_with("Return"))
returns <- returns %>% select(-starts_with("Death"))

# Combine the deaths and returns datasets
combined <- deaths %>%
  left_join(select(returns, URL, Name.Alias, Time, Return), by = c("URL", "Name.Alias", "Time"))

#Get rid of some characters with empty names
combined_filtered <- combined %>%
  filter(Name.Alias != "")

#Get a df of each character and how many times they died
avg_deaths <- combined_filtered %>%
  group_by(Name.Alias) %>%
  summarise(total_deaths = sum(Death == "YES", na.rm = TRUE))

#Gets the mean of all deaths
average_num_deaths <- mean(avg_deaths$total_deaths, na.rm = TRUE)

average_num_deaths
```
## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Blake Nelson

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.


  "But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death."
    
### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

```{r}
# Count the number of unique characters who had at least one "YES" in the return column
characters_with_return <- combined_filtered %>%
  filter(Death == "YES") %>%
  distinct(Name.Alias)  # Count the unique characters

# Count the total number of unique characters
total_characters <- combined_filtered %>%
  distinct(Name.Alias)

# Calculate the percentage
percentage_with_return <- nrow(characters_with_return) / nrow(total_characters) * 100

percentage_with_return
```

```{r}
# Count the number of unique characters who returned after more than one death
characters_with_multiple_returns <- combined_filtered %>%
  filter(Death == "YES") %>%
  group_by(Name.Alias) %>% 
  summarise(num_returns = sum(Return == "YES", na.rm = TRUE)) %>%
  filter(num_returns > 1) %>%
  distinct(Name.Alias)  # Count the unique characters

# Calculate the percentage
percentage_with_multiple_returns <- nrow(characters_with_multiple_returns) / nrow(total_characters) * 100

percentage_with_multiple_returns
```
### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

  There is not a 2 in 3 percent chance that a character returns after dying. it is more like 2 in 5 chance or 40%. There is also only around 5% chance that a character returns more than once.  The fact is clearly not true.

## Ishan Patel
```{r}
#Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team. Let’s put it this way: If you fall from four or five stories up, there’s a 50 percent chance you die. Getting a membership card in the Avengers is roughly like jumping off a four-story building.

# Calculate the number of Avengers who have died at least once

#Commented because it does not work

# avengers_died_at_least_once <- av %>%
#   filter(Reduce(`|`, across(starts_with("Death"), ~. == "YES")))
# 
# # Calculate the percentage of Avengers who have died at least once
# percentage_died_at_least_once <- nrow(avengers_died_at_least_once) / nrow(avengers) * 100
# 
# percentage_died_at_least_once

# The statement is true.
```

## Hazer Becic


