---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
#imports
library(dplyr)
library(tidyverse)
```

```{r}
# Define function to parse numbers
parse_number <- function(x) as.numeric(gsub("\\D", "", x))

# Replace columns for deaths
av <- av %>%
  pivot_longer(cols = starts_with("Death"), 
               names_to = "Time_Death", 
               values_to = "Death") %>%
  mutate(Time_Death = parse_number(gsub("Death", "", Time_Death)),
         Death = ifelse(Death == "YES", "yes", ifelse(Death == "NO", "no", ""))) %>%
  select(-Time_Death)  # Remove the temporary column used for pivoting

head(av)
```

Similarly, deal with the returns of characters.

```{r}
# Replace columns for returns
av <- av %>%
  pivot_longer(cols = starts_with("Return"), 
               names_to = "Time_Return", 
               values_to = "Return") %>%
  mutate(Time_Return = parse_number(gsub("Return", "", Time_Return)),
         Return = ifelse(Return == "YES", "yes", ifelse(Return == "NO", "no", ""))) %>%
  select(-Time_Return)  # Remove the temporary column used for pivoting

head(av)
```
Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
# Calculate the total number of deaths for each Avenger
av <- av %>%
  mutate(Total_Deaths = rowSums(select(., starts_with("Death")) == "yes", na.rm = TRUE))

# Calculate the average number of deaths
average_deaths <- mean(av$Total_Deaths, na.rm = TRUE)

# Print the average number of deaths
print(average_deaths)
```
## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Blake Nelson


## Ishan Patel
```{r}
#Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team. Let’s put it this way: If you fall from four or five stories up, there’s a 50 percent chance you die. Getting a membership card in the Avengers is roughly like jumping off a four-story building.

# Calculate the number of Avengers who have died at least once
avengers_died_at_least_once <- avengers %>%
  filter(Reduce(`|`, across(starts_with("Death"), ~. == "YES")))

# Calculate the percentage of Avengers who have died at least once
percentage_died_at_least_once <- nrow(avengers_died_at_least_once) / nrow(avengers) * 100

percentage_died_at_least_once

# The statement is true.
```

## Hazer Becic
